---
title: "Lecture 12: Gelman Hill Ch 5"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse) 
library(arm)
library(knitr)
```


# Logistic Regression

When the research goal is model or predict a binary outcome, modifications need to be made to the linear model framework that we have previously used.

Before, we had the following specification:

$$y_i = X_i \underline{\beta} + \varepsilon_i,$$
where $\varepsilon \sim N(0,\sigma^2)$ or equivalently $y_i \sim N(X\underline{\beta}, \sigma^2)$.

\vfill


\vfill
\vfill

### Logistic Regression with a single predictor

One of the classical datasets used for regression involved o-rings on the challenger space shuttle.

[https://www.space.com/31732-space-shuttle-challenger-disaster-explained-infographic.html](https://www.space.com/31732-space-shuttle-challenger-disaster-explained-infographic.html)

From a NASA press release...

>> On January 28, 1986, as the Space Shuttle Challenger broke up over the Atlantic Ocean 73 seconds into its flight, Allan McDonald looked on in shock -- despite the fact that the night before, he had refused to sign the launch recommendation over safety concerns. 

>> McDonald, the director of the Space Shuttle Solid Rocket Motor Project for the engineering contractor Morton Thiokol, was concerned that below-freezing temperatures might impact the integrity of the solid rockets' O-rings. 


\newpage

The `orings` dataset is available in the `faraway` package in R. The data contains the number of o-rings that failed on previous launches (out of a total of 6) along with the temperature on the day of the launch.
```{r, message = FALSE}
library(faraway)
data(orings)
orings_updated <- orings %>% mutate(non_damage = 6 - damage, 
                                    failure_proportion = damage/6)
kable(orings_updated, digits = 2)
```

The temperature on the morning of the crash was 31F. We are going to build a model to explain the impact of temperature on the failure probability of the o-rings.

\vfill

\newpage
```{r}
orings_updated %>% ggplot(aes(y=failure_proportion, x = temp)) + geom_point() +
  geom_smooth(method = 'loess', color = 'red') + 
  ggtitle('O-Ring Failure proportion as a function of launch temperature') +
  ylab('Failure proportion \n (out of 6 o-rings)') + xlab('Launch Temperature (F)') +
  theme_bw() + geom_smooth(method = 'lm')
```

\vfill

Using the `glm` function we can fit a logistic regression model. Note in this case the response is a data frame with successes and failures.


```{r}
challenger_glm <- glm(cbind(damage, non_damage) ~ temp,data = orings_updated, family = binomial)
display(challenger_glm)
```

\newpage
```{r}
temp_tibble <- tibble(temp = seq(30,85, by = .1)) %>% 
  mutate(pred = predict(challenger_glm, ., type = 'response'))

orings_updated %>% ggplot(aes(y=failure_proportion, x = temp)) + geom_point() +
  geom_smooth(method = 'loess', color = 'red') + 
  ggtitle('O-Ring Failure proportion as a function of launch temperature') +
  ylab('Failure proportion \n (out of 6 o-rings)') + xlab('Launch Temperature (F)') +
  geom_line(aes(y = pred, x=temp), data = temp_tibble, inherit.aes = F) +
  theme_bw() 
```

